\chapter{Improved Melanoma Detection using Asymmetry}

\section{Introduction}

Melanoma is a type of malignant skin cancer that accounts for a significant proportion of cancer-related deaths around the world. In 2018 there were approximately 2,353 per 100,000 deaths in the United Kingdom (UK)\cite{UK2019}. Early detection is critical for improving the diagnosis and survival of patients. However, existing approaches including clinical examinations and dermoscopy, have limitations in terms of accuracy and cost-effectiveness\cite{Takiddin2021}. Machine learning approaches have beaten dermatologists in terms of accuracy\cite{Andre2017}. However, these approaches lack explainability implementing such techniques difficult for clinical environments\cite{Jiang2017}. One concern is the production of realistic, but incorrect results\cite{Ghorbani2019}. Another is the use of parallel processes, which describes the creation of an answer with little to no explanation. In this paper, we propose a combined asymmetry approach using shape, colour, and texture analysis alongside a detailed comparison. The technique itself can be used in conjunction with ABCD rules (Asymmetry, border, colour, dermoscopic features).

%------------------------------------------------


\section{Related Works}

% Quick summary of TDS, ABCD rules, bi-fold and general issues
Asymmetry analysis is a fundamental component in the early detection of melanoma because it often exhibits asymmetric shapes\cite{}. Bi-fold is a diagnostic procedure designed to recognize melanoma by drawing a line down the middle of the skin lesion and comparing the two halves to confirm whether the sides match (considering the difference in shape, colour, and texture). Using this horizontally and vertically calculates whether the skin lesion is possibly malignant with a score between 0 and 2. Calculating with Total Dermoscopy Score (TDS) alongside the other ABCD rules including asymmetry, border, colour, and diameter calculates the likelihood of malignancy. Dermatologists frequently use bi-fold due to its simplicity, but it can be subjective to the original observer and time-consuming when managing large numbers of skin lesions. Therefore, automating techniques is beneficial to clinicians and can improve the objectivity of results.

% Measuring asymmetry shape
Ihab S. Zaqout \cite{Zaqout2016} describes a technique using the centroid and rotation of the skin lesion using moments of inertia. By Folding the skin lesion on both vertical and horizontal axes subtracting the opposite half. Pixels that cannot subtract are summed and compared with a threshold considering the skin lesion asymmetrical if the combined sum is more than the threshold.

% Measuring asymmetry colour
Kasmi and Mokrani \cite{Kasmi2016a} create a grid of 20 by 20 pixels from the skin lesion image and convert it into the LAB colour space. They then compare the average colour of each block with a perpendicular block (vertical and horizontal axes) using the three-dimensional Euclidean luminance distance, a-axis, and b-axis. If more than half of the colour comparisons exceed the threshold, they consider that axis to be colour asymmetrical. They ignore blocks that have no symmetrical pair. Finally, they calculate luminance separately to prevent brightness problems. This technique achieves an accuracy of 94\% with a private dataset.

% Measuring asymmetry texture
Ali \cite{Ali2020a} uses SIFT-based similarity and projection profiles to measure similarities in texture. SIFT is scale-invariant and helpful for texture components with varying texture quality. First, they split the skin lesion vertically and horizontally across the centre into four halves, compare texture components on the symmetrical halves, and measure similarity. Lastly, they generate histograms for the projection profile in the x and y directions. These results train a decision tree and achieve an 80\% accuracy of the ISIC 2018 with 204 images privately annotated for ABCD rules.

% Summary of techniques and mention of the proposed approach
Prior studies have introduced techniques that measure distinct aspects of asymmetry, such as Ihab S. Zaqout \cite{Zaqout2016} measurement of shape, Kasmi and Mokrani \cite{Kasmi2016a} measurement of colour, and  Ali \cite{Ali2020a} measurement of texture. The new approach seeks to combine the following approaches into a more comprehensive analysis of asymmetry that takes into account multiple features of the skin lesion. The proposed novel technique updates colour measurement to improve accuracy using superpixels and an SVM model.

\section{Asymmetry Analysis Approach}
% Describe the asymmetry technique and the changes that have been made to it.
The proposed method has many similarities to Kasmi and Mokrani \cite{Kasmi2016a} colour comparison technique, except it is updated to improve accuracy using a number of techniques. Superpixels are k-means colour extraction techniques to separate areas of an image into their associated areas of colour by applying a soft border around an edge. The premise is the original method of splitting the lesion into a 20 by 20 grid without any flexibility splits areas of interest in half, making comparisons less adequate. Using superpixels allows for a softer border which improves colour separation and accuracy.

To initiate the classification of skin lesions a technique called bi-fold is applied involving folding the skin lesion in half vertically and horizontally and a comparison of their respective dimensions. While the original technique was designed only to assess the lesions' shape, it's been utilized to account for colour and texture as well. The centre and orientation are determined by calculating its moments, where the centre is (m10 / m00, m01 / m00) and phi is 0.5tan(2m11)/(m20 -m02). 

Next, the lesion is partitioned into a 20 by 20 grid centred on the mentioned centre point, and the average of each region is computed. This is followed by finding the matching region on the perpendicular area from the centre of the skin lesion and comparing the colour distance between the two. Distance is measured using the LAB colour space and a 2d Euclidean distance of A and B, removing L (luminosity) to eliminate light variation. Once compared, all compared regions are obtained, and they are plotted onto a graph. If over half of the values are above a threshold of 8, then the lesion is asymmetrical.

% Scrutinize the threshold method and show the graph that the data can't be split with only the threshold 
The weakest point in the technique is the use of a threshold, if more than half of the subtracted skin lesions are 8 it is considered asymmetrical. To demonstrate why this method is not suitable, look at the graph below using the PH^2 dataset. Essentially, the symmetrical skin lesion has a smaller area and the asymmetrical lesion has a larger area, but both remain in the same zone and therefore splitting the data only using a threshold has poor results. Furthermore, there are a lot of outlying data and the threshold does not adjust according to these values. See the graph below:

% Demonstrating the changes including superpixels and graphs, etc

