\chapter{NHS Data Suitability for Melanoma Detection Using Machine Learning Algorithms}

\section{Introduction}
The use of machine learning algorithms for the detection of melanoma is a promising and evolving field with detection accuracies often beating that of a dermotologists\cite{Andre2017}. However, the effectiveness of such depends heavily on the quality of the datasets used to develop them\cite{Tae2019}. The goal of this chapter is to describe the data extraction process from the National Health Service (NHS) and highlight biases, preprocessing, and other potential issues involved in the training of machine learning algorithms for the detection of melanoma.


\section{Background}
%Describe the requirements, such as using macroscopic images over dermoscopic.

There are requirements for the project including the use of macroscopic images instead of dermoscopic images. Macroscopic is described as viewing with a naked eye or by taking a standard picture with no special lenses. Dermoscopic is when you use a specialized tool called a dermoscope.

Dermoscopy improves the diagnostic accuracy of dermatologists for melanoma when compared with macroscopic examination\cite{Wolner2017} and is widely considered superior\cite{Thiers2009}. Dermoscopic images provide a detailed visualization of patterns and structures on the surface of the skin lesion that might not be visible to the naked eye\cite{Thiers2009}. Some of these structures are pigment network, asymmetry, irregular borders, and other features that support in the differentiation between benign and malignant lesions\cite{Thiers2009}. 

Another example, shows the diagnosis for BCC was 91\% when using dermoscopy, compared to 57\% when using close-up images\cite{Dascalu2022}. Similarly, the sensitivity for SCC was 77\% with dermoscopy, compared to 70\% with close-up images\cite{Dascalu2022}. These findings highlight the superior diagnostic performance of dermoscopy compared to macroscopic.

Dermoscopic examination is superior to macroscopic examination, however, the projects requirements specify using macroscopic. The logic behind this is that general practitioners are unlikely to recognize dermoscopic features, so there is no need to supply them with dermoscopes.  This appears to be consistent with an authors findings showing that 92\% of dermatologists correctly recognize at least four of size types of melanoma. In contrast, only 38\% of non-dermatologists were able to recognize the same number of melanomas\cite{Tae2019}. Therefore, `the dataset' is created with macroscopic images for examination.

Considerations to clean the data removing hair and specular reflection to improve classification accuracy. This chapter will discuss the data transformation of NHS macroscopic images, including augmentation techniques to remove lighting, hair and other anomalous data from the images. All of which will support in improving the accuracy when classifying.

Ideally the metadata included would include Filename, Tags, Gender, DOB, Department, Consent, Diagnosis, Date Photographed. However, there has been difficulty accessing this data, and therefore only the diagnosis of the skin lesions are available.

The goal of this chapter is to measure the usefulness of NHS data, highlighting potential issues or biases in the data.

\section{Other Public Datasets}
Overall there are a range of public datasets including MEDLINE, PH2, ISIC, and others making a total of 21 open access datasets containing 106,950 skin lesion images\cite{Wen2022}. Out of these datasets only the PH2 dataset has publicly accessible data on the ABCD rules with a total of 200 images.


\section{Data Biases}
The use of datasets is fundamental to the development and evaluation of machine learning algorithms, and the accuracy and effectiveness heavily weigh on the quality of the data used. Biases can arise from data collection procedures and pre-processing techniques. Not considering possible biases greatly affects machine learning algorithms using them and their effectiveness. Furthermore, careful consideration is essential to ensure the accuracy and reliability of the conclusions proposed in this document. Failure to consider all these factors could result in skewered conclusions that could undermine the validity of findings. For these reasons, it is essential to carefully identify and evaluate data before using and testing it.

The NHS datasets contain a wealth of information that can be utilized. However, some biases need consideration before creating a dataset. These biases include:
\begin{enumerate}
    \item  The diagnostic procedure dismisses skin lesions without recognizably suspicious features and do not reach the phase that photographs were captured. As such, there is a lack of typical benign skin lesions within the dataset, and most have some undesirable features.
    \item Dermatologists have diagnosed the large majority of skin lesions which have varying accuracy depending on their experience. There is no way of knowing how accurate this data is.
    \item Dermatologists could diagnose during an in-person examination where patients can be asked questions in real-time and further tests can be made involving touch. Otherwise, dermatologists diagnose using previously saved images, which might be less accurate because it is lacking the insight that an in-person examination would provide.
    \item Some skin lesions within the dataset are lacking metadata including their diagnosis. Such image samples should be avoided.
    \item Diagnoses of skin lesions are written in plain text including questions marks where there is some uncertainty and the possibility of multiple diagnoses. Only diagnoses that are certain of their findings are used.
    \item Photographs of the skin lesions may be captured on different body parts such as hands, legs, face, and others. Most pre-processing methods are designed to differentiate between skin and skin lesions, so it is important to avoid using these images. Otherwise, new pre-processing methods will have to be made and tested.
    \item Seborrhoeic keratosis (SK) have similar features to that seen in malignant skin lesions. Therefore, there might be skin lesions diagnosed as melanoma that are SK. Furthermore, because of its similarity there are potentially many SK images. It will be vital to separate these.
\end{enumerate}

Varied accuracy of

\section{Image Criteria}

The dataset includes skin lesions of Malignant Melanoma (MM), Seborrhoeic keratosis (SK), Atypical Naevi (AN), Typical Naevi (TN), Squamous Cell Carcinoma (SCC), Basal Cell Carcinoma (BCC). 

Considering the data biases described in the previous section we can begin to piece together criteria that describe which images are appropriate for the creation of the 'dataset'. Firstly, the most glaring problems are the diagnoses, because some images are missing diagnostic data and some have a variety of diagnoses, the scenario in which the diagnoses were captured is not mentioned, 
SK is potentially misdiagnosed as malignant, and the dermatologist that originally diagnosed are not mentioned. There are many ways to combat these problem. For example during the creation of the PH2 dataset\cite{mendonca2013}, several dermatologists were asked to diagnose a skin lesion with appropriate metadata relating to the ABCD rules. If most dermatologists agree it is included in the study, otherwise it is removed. The goal of this is to minimize the incorrect data within the dataset. Alternatively, you might argue that removing less adequate records fails to prove whether the developed algorithms work in a real medical environment. Therefore, other methods were explored to validate the data.

%Very Unlikely we'll be getting histopathology data
In this project, we do not have the resources to re-diagnose skin lesions, so instead the focus is on gathering diagnostic results from the histopathology department. These results are more accurate than dermatologists\cite{Morton1998} and counter the mentioned biases. Comparing this data with the diagnoses from the dermatology department provides insight into dermatology accuracy. Furthermore, histopathology serves as a ground-truth for training the algorithms and testing their capabilities.

As mentioned in the data biases section the skin lesion images are taken under various different conditions including angles, lighting, and distance from the skin lesion. While the variety of conditions will decrease the accuracy of results and hinder the detection of dermoscopic features, it is a requirement of the project. 

These variations in image quality would decrease the accuracy of results. For this study the focus is on skin lesions captured using dermoscopes, avoiding the cases where skin lesions are captured under less-adequte conditions.

Use ANOVA to check whether features of Atypical and typical moles are similar.

%Data extraction Techniques


%Show transforming the images into the correct size


%Specular reflection removal
\section{Skin Lesion Augmentation}
One of the main challenges in melanoma detection is the visual similarity normal and infected regions. Others are the presence of artefacts such as bubbles, hair and clinical marks\cite{Albahli2020}. These factors lead to low accuracy rates in traditional approaches. However, segmentation techniques can help overcome these challenges by removing these areas and isolating the melanoma from the rest of the image.

Skin lesion augmentation is especially vital because of the use of macroscopic images instead of dermoscopic images. This means there are various artefacts including hair, specular reflections, rulers, varying sizes, and shapes of the skin lesion. All of these can obscure the skin lesion and affect the accuracy of segmentation\cite{Unver2019} and in effect feature detection.

By augmenting the skin lesion images using specular reflection removal and hair removal, the accuracy of feature classification methods can be improved\cite{kasmi2023}. 

\subsection{Hair Removal}
Hair artificats in images can interfere with the recognition of handcrafted features and affect the performance of deep learning algorihms in melanoma detection\cite{kasmi2023}. Applying morphological operations such as image sharpening and segmentation technqiues can remove hair artifacts from dermoscopic images\cite{kasmi2023}.

Dull-Razor is an algorithm developed by Lee et al\cite{Lee1997} and is frequently implemented with 

Sharp-Razor\cite{kasmi2023} is a technique for detecting hair and ruler marks to remove them from  images. This uses a multiple filter approach including grayscale plane modification, hair enhancement, segmentation using tri-directional gradients, and multiple filters for hair of varying widths. This technique is shown to outperform exisiting methods.

\subsection{Specular Removal}
Specular reflection removal techniques have been shown to be effective in improving the accuracy of melanoma detetion\cite{Shen2009}. A technique was proposed utilizing a partial differential equation to iteratively erode the specular compnonent, removing the specular reflection\cite{Shen2009}.
